The \vspec{} Collection: A suite of utilities to model spectroscopic phase curves of 3D exoplanet atmospheres in the presence of stellar variability.

### 2024-08-02 src/tex/ms.tex
Add pypsg code listing

\author[0000-0002-1570-2203]{Ted M Johnson}
\affiliation{Southeastern Universities Research Association, 1201 New York Avenue NW, Suite 430, Washington, DC 20055, USA}
\affiliation{Center for Research and Exploration in Space Science and Technology, NASA Goddard Space Flight Center, 8800 Greenbelt Road, Greenbelt, MD 20771, USA}
\affiliation{NASA Goddard Space Flight Center \\
8800 Greenbelt Rd \\
Greenbelt, MD 20771, USA}
\affiliation{Nevada Center for Astrophysics, University of Nevada, Las Vegas, 4505 South Maryland Parkway, Las Vegas, NV 89154, USA}
\affiliation{Department of Physics and Astronomy, University of Nevada, Las Vegas, 4505 South Maryland Parkway, Las Vegas, NV 89154, USA}

@@

\vspec{} configurations are designed to minimize human errors by providing two equivalent formats. The first, a file written in YAML, is optimized for
human readability. For example, the \texttt{system} section, which describes the relationship between the observed planetary system and the observer, could
be written:
\begin{lstlisting}[label={ls:yaml}]
system:
    distance: 12.4 pc
    inclination: 89.7 deg
    phase_of_periasteron: 0 deg
\end{lstlisting}
Note that the units of each of these parameters is included in a human-readable way. Internally, \vspec{} casts each of these parameters to an Astropy
\texttt{Quantity} instance, so that the user can input any desired value and unit combination so long as the physical type is correct -- i.e. it would be
equivalent to write \texttt{3.83e17 m} as the value for the \texttt{distance} parameter.

The second input method available is to directly initialize the Python object that \vspec{} uses internally to store its model parameters.
The top-level object, \texttt{InternalParameters}, is structured to mirror the YAML input file, with one argument per YAML section. The
parameters in Listing \ref{ls:yaml} would be written:
\begin{lstlisting}[language=Python]
from astropy import units as u
from VSPEC.params import *
params = InternalParameters(
    ... # Other arguments
    system=SystemParameters(
        distance=12.4*u.pc,
        inclination=89.7*u.deg,
        phase_of_pariasteron=0*u.deg
    )
)
\end{lstlisting}
This is convenient for producing configurations programmatically, for example for producing a grid of model phase curves.

@@

\vspec{} uses a Python interface for the PSG API system called \texttt{pypsg}, which is a general stand-alone library built for any Python-based API call to PSG. \texttt{pypsg} draws inspiration from object-relational mapping in frameworks such as Django\footnote{\url{https://www.djangoproject.com/}} to encode a PSG configuration file as a native Python object.Fields of a \texttt{pypsg} data model represent one or more lines of a configuration file, and the two representations are interchangable without loss of information. Upon creation of this configuration as a \texttt{PyConfig} object, a user can create an \texttt{APICall} instance which handles sending a request to PSG via Python's \texttt{requests}\footnote{\url{https://github.com/psf/requests}} package. This dedicated caller is important in that it reads a user's API key from a known location on disk, removing the possibility that the key
could be committed to a public repository.

@@

>>>>

% \author[0000-0002-1570-2203]{Ted M Johnson}
\author{Ted M Johnson}
\affiliation{Southeastern Universities Research Association, 1201 New York Avenue NW, Suite 430, Washington, DC 20055, USA}
\affiliation{Center for Research and Exploration in Space Science and Technology, NASA Goddard Space Flight Center, 8800 Greenbelt Road, Greenbelt, MD 20771, USA}
\affiliation{NASA Goddard Space Flight Center \\
8800 Greenbelt Rd \\
Greenbelt, MD 20771, USA}
\affiliation{Nevada Center for Astrophysics, University of Nevada, Las Vegas, 4505 South Maryland Parkway, Las Vegas, NV 89154, USA}
\affiliation{Department of Physics and Astronomy, University of Nevada, Las Vegas, 4505 South Maryland Parkway, Las Vegas, NV 89154, USA}

@@

\vspec{} configurations are designed to minimize human errors by providing two equivalent formats. The first, a file written in YAML, is optimized for
human readability. For example, the \texttt{system} section, which describes the relationship between the observed planetary system and the observer, could
be written:
\begin{lstlisting}[label={ls:yaml},caption={\vspec{} YAML Configuration}]
system:
    distance: 12.4 pc
    inclination: 89.7 deg
    phase_of_periasteron: 0 deg
\end{lstlisting}
Note that the units of each of these parameters is included in a human-readable way. Internally, \vspec{} casts each of these parameters to an Astropy
\texttt{Quantity} instance, so that the user can input any desired value and unit combination so long as the physical type is correct -- i.e. it would be
equivalent to write \texttt{3.83e17 m} as the value for the \texttt{distance} parameter.

The second input method available is to directly initialize the Python object that \vspec{} uses internally to store its model parameters.
The top-level object, \texttt{InternalParameters}, is structured to mirror the YAML input file, with one argument per YAML section. The
parameters in Listing \ref{ls:yaml} would be written:
\begin{lstlisting}[language=Python, caption={\vspec{} Python configuration}]
from astropy import units as u
from VSPEC.params import *
params = InternalParameters(
    ... # Other arguments
    system=SystemParameters(
        distance=12.4*u.pc,
        inclination=89.7*u.deg,
        phase_of_pariasteron=0*u.deg
    )
)
\end{lstlisting}
This is convenient for producing configurations programmatically, for example for producing a grid of model phase curves.

@@

\vspec{} uses a Python interface for the PSG API system called \texttt{pypsg}, which is a general stand-alone library built for any Python-based API call to PSG. \texttt{pypsg} draws inspiration from object-relational mapping in frameworks such as Django\footnote{\url{https://www.djangoproject.com/}} to encode a PSG configuration file as a native Python object. Fields of a \texttt{pypsg} data model represent one or more lines of a configuration file, and the two representations are interchangable without loss of information. Upon creation of this configuration as a \texttt{PyConfig} object, a user can create an \texttt{APICall} instance which handles sending a request to PSG via Python's \texttt{requests}\footnote{\url{https://github.com/psf/requests}} package. This dedicated caller is important in that it reads a user's API key from a known location on disk, removing the possibility that the key could be committed to a public repository.

@@

Listing \ref{ls:pypsg} shows a minimal example that uses \texttt{pypsg} to make a call to PSG/GlobES. Here \texttt{pypsg} converts a GCM output from the netCDF format to a \texttt{PyGCM}, which, along with data model classes like \texttt{Target} and \texttt{Observatory}, is used as an argument to construct a \texttt{PyConfig} class instance. This configuration object is then passed to an \texttt{APICall} instance. When the API call is made, the contents of the \texttt{PyConfig} are included in the body of the POST request in the \texttt{file} field. PSG reads this the same as it would if we had written those contents to a file \texttt{config.txt} and passed the \texttt{--data-urlencode file@config.txt} option to \texttt{cURL}, as are the instructions in the PSG handbook \citep[pp. 162-163]{villanueva2022}.

\begin{lstlisting}[
    basicstyle=\footnotesize,language=Python,label={ls:pypsg},caption={\texttt{pypsg} Example}
]
    import pypsg
    from astropy import units as u
    from netCDF4 import Dataset

    # This next line only needs to be run once per user.
    # These settings are saved to `~/.pypsg/settings.json`
    pypsg.settings.save_setting(
        api_key='my_api_key',
        url='https://psg.gsfc.nasa.gov/api.php'
    )

    gcm_data = Dataset('path/to/gcm.nc')

    cfg = pypsg.cfg.PyConfig(
        target=pypsg.cfg.Target(
            object='Exoplanet',
            name='proxima-cen-b'
            ...
        ),
        geometry=pypsg.cfg.Observatory(...),
        generator=pypsg.cfg.Generator(...),
        telescope=pypsg.cfg.SingleTelescope(...)
        noise=pypsg.cfg.CCD(...)
        gcm=pypsg.globes.waccm_to_pygcm(
            gcm_data,
            itime=0,
            molecules=['H2O','CO2'],
            aerosols=['Water','WaterIce'],
            mean_molecular_mass=48.0
        )

    )

    caller = pypsg.APICall(
        cfg=cfg,
        app='globes',
        output_type='rad'
    )
    response = caller()

    print(response.rad['Wave/freq','proxima-cen-b'])

    -------------------

    Wave/freq  proxima-cen-b
        um      W / (um m2) 
    ----------- -------------
            1.0   1.81314e-20
           1.02    3.1926e-20
         1.0404   4.95983e-19
       1.061208    5.0892e-19
     1.08243216   3.79632e-19
            ...           ...
     16.3163953   7.66301e-20
     16.6427232   7.66245e-20
    16.97557767   7.54221e-20
    17.31508922   7.49258e-20
    17.66139101   7.35507e-20
    Length = 146 rows

\end{lstlisting}
